---
title: "Reproducible Research: Peer Assessment 1"
author: "Mike Liu"
date: "6/18/2020"
output: 
  html_document:
    keep_md: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
## Loading and preprocessing the data
1. Reading in Dataset
```{r}
df = read.csv("activity.csv")
df$date = as.Date(df$date)
```

2. Histogram of total numbers of steps taken each day
```{r, echo=FALSE}
tot_steps_per_day <- aggregate(df$steps, by=list(df$date), FUN=sum)
hist(tot_steps_per_day$x, main="Histogram of Total Number of Steps per Day", xlab="Sum of Steps Per Day")
```
## What is mean total number of steps taken per day?
3. Mean and median number of steps taken each day:
```{r}
print(paste("Mean steps taken each day:", mean(tot_steps_per_day$x[!is.na(tot_steps_per_day$x)])))
print(paste("Median steps taken each day:", median(tot_steps_per_day$x[!is.na(tot_steps_per_day$x)])))
```
## What is the average daily activity pattern?
4. Time-series plot of average number of steps taken
```{r}
plotdf = df[!is.na(df$steps),]
avg_steps_per_day <- aggregate(plotdf$steps, by=list(plotdf$date), FUN=mean)
plot(x=avg_steps_per_day$Group.1, y=avg_steps_per_day$x, 
     main = "Avg Steps per Day Over Time",
     xlab = "Date", ylab = "Avg Steps",
     type="l")
```

5. 5-min interval that contains the max number of steps
```{r}
five_min_int_steps <- aggregate(plotdf$steps, by=list(plotdf$interval), FUN=mean)
max_int <- with(five_min_int_steps, Group.1[x=max(x)])
print(paste("5-min interval that contains the max number of steps:", max_int))
```

## Imputing missing values
6. Code to describe and show a strategy for imputing missing data
```{r}
library(dplyr)
# Using a interval-mean imputation strategy for missing steps
df_imp <- df %>% group_by(interval) %>% mutate(steps = ifelse(is.na(steps), mean(steps, na.rm=TRUE),steps))
print(head(df_imp))
```

7. Histogram of total number of steps taken each day after missing values are imputed
```{r}
imp_tot_steps_per_day <- aggregate(df_imp$steps, by=list(df_imp$date), FUN=sum)
hist(imp_tot_steps_per_day$x, main="Histogram of Total Number of Steps per Day", xlab="Sum of Steps Per Day")
```

## Are there differences in activity patterns between weekdays and weekends?
8. Panel plot comparing avg number of steps taken per 5-min inteval across weekdays and weekends
```{r}
# Add in boolean mask for weekend or not
df_imp$Weekday <- ifelse(weekdays(df_imp$date) %in% c("Saturday", "Sunday"), FALSE, TRUE)

# Use aggregate to aggregate weekday and weekend steps
avg_steps_weekday <- with(df_imp[df_imp$Weekday,], aggregate(steps, by=list(interval), FUN=mean))
avg_steps_weekend <- with(df_imp[!df_imp$Weekday,], aggregate(steps, by=list(interval), FUN=mean))

# Set up panels, plot line charts for weekdays and weekends
par(mfrow = c(1,2))
plot(x=avg_steps_weekday$Group.1, 
     y=avg_steps_weekday$x, 
     xlab = "5-Min Interval", ylab = "Avg Steps",
     main="Avg Steps per Interval, Weekdays", type="l")
plot(x=avg_steps_weekend$Group.1, 
     y=avg_steps_weekend$x, 
     xlab = "5-Min Interval", ylab = "Avg Steps",
     main="Avg Steps per Interval, Weekends", type="l")

```


